On The Difficulty of Computing Non-Commutative Signature Gröbner Bases

### 2022-01-26 article_issac.tex
Reprise s -> ss dans l'exemple du langage non algébrique
Traduction de la preuve dans le cas non-compatible.

\theoremstyle{definition}
\newtheorem{defn}[lem]{Definition}

@@

@@

@@

\section{Finiteness}

>>>>

\theoremstyle{definition}
\newtheorem{defn}[lem]{Definition}
\newtheorem*{fact}{Fact}

@@

\newcommand{\sss}{\mathfrak{s}}

@@

\newcommand{\GS}{G^{[\Sigma]}}

@@

\section{Fairness}

\section{Compatibility}

In this Section,
we provide an example
of a set of polynomials
such that 
it has a finite GB,
its module of
syzygies is generated
by one syzygy plus
the principal syzygies,
but there is
a non-compatible order
such that it
has no finite signature GB.

We set $X = \{t, x, a, b \}$.
We take as order on 
$\left\langle X \right\rangle$ the 
graded lexicographic order
with $a < b < t < x$.

Let us define:
\begin{align*}
	g_1 &:= xt - ta \\
	g_2 &:= xt - tb \\
	g_3 &:= b - a
\end{align*}
One shows easily
thanks to Buchberger's criterion
that $(g_1,g_2,g_3)$
is a GB of the ideal $I$ it spans.

We define the following
fair but non-compatible order
on $\MM \left( \mathscr{F}_3 \right):$
  $(l, i, r) \leq (l', j, r')$ if one of the following
  conditions is satisfied:
\begin{itemize}
	\item $\deg{l\LM(g_i)r} < \deg{l'\LM(g_j)r'}$
	\item $\deg{l\LM(g_i)r} = \deg{l'\LM(g_j)r'}$ and $i < j$.
	\item $\deg{l\LM(g_i)r} = \deg{l'\LM(g_j)r'}$ and $i = j$ and $\deg{r} < \deg{r'}$.
	\item $\deg{l\LM(g_i)r} = \deg{l'\LM(g_j)r'}$ and $i = j$ and $\deg{r} = \deg{r'}$ and $l\LM(g_i)r \geq l'\LM(g_j)r'$.
	% Penser à xxt xta taa tab xtb pour voir que le signe doit être dans l'autre sens.
\end{itemize}

We define the 
degree of an element of 
$\MM \left( \mathscr{F}_3 \right)$
by $$\deg{(l, \varepsilon_i, r)} = \deg{l \LM(g_i) r}$$.
This extends to $\mathscr{F}_3$
by stating that the 
degree of 
an element is the maximal
degree of the monomials
of its terms.

\begin{lem}
	In this context
	$I^\Sigma$ has no
	finite signature GB.
\end{lem}
\begin{proof}
We proceed by contradiction.
Let us assume that $G^{[\Sigma]}$ is such 
a basis.
Using inter-reduction,
we may assume that
$G^{[\Sigma]}$
is reduced.

	\begin{fact}
		For any $(f,\sigma)$ in $G_f$, $\deg{\LM(f)} = \deg{\LM(\sigma)}$.
	\end{fact}
	\begin{proof}
	If it were not the case,
	 thanks to
	homogeneity,
	the terms of 
	degree  $\deg{\LM(\sigma)}$ in $\sigma$ would correspond to
	a syzygy, which contradicts our hypothesis
	that $(f,\sigma)$ is
	irreducible.
	\end{proof}

Now, for any $n \in \N$,
	$$\left(ta^nb - ta^{n+1}, - (x^n, 2, 1)
		- \sum_{i = 0}^{n - 1}(x^i, 1, a^{n - 1 - i}b) + 
		\sum_{i = 0}^{n} (x^i, 1, a^{n - i})
		  \right)$$
	is in $I^{[\Sigma]}.$
Hence, it has to be
$\sss$-reducible
by some 
 $(f_n, \sigma_n)$ of $G^{[\Sigma]}$.
 There is then some
 $l, r \in \KX$ such that:
	\begin{align*}
		l\LM(f_n)r = ta^nb, \\
		l \LM(\sigma_n) r \leq (x^n, 2, 1).
	\end{align*}

We take some $n$
such that 
the degree of $ta^nb$
is strictly bigger
than any $\LM$
of a polynomial
in $\GS$. 
	Therefore, $\LM(f_n)$ is one of the following: $a^k$, $a^kb$, or $ta^k$,
	for some $k.$
Two cases are then possible:	
	\begin{itemize}
		\item $\LM(f_n)$ 
		has no $t$. 
		In that case,$\LM(f_n)$  can only
be reached using 
a term of the form
$(u, \varepsilon_3, v)$ 
in $\sigma_n$, with $\deg (u, \varepsilon_3, v) = \deg \LM(f_n)$,
and this is a contradiction
since:
		$$(lu, \varepsilon_3, vr) \leq l \LM(\sigma_n) r \leq (x^n, \varepsilon_2, 1),$$
		with $\deg (lu, 3, vr) = \deg (x^n, 2, 1)$.

		\item Otherwise, $\LM(f_n) = ta^k$.
		This is actually
not possible:
$ta^k$ is divisible
by none of the $\LM(g_i)$'s.
	\end{itemize}
Consequently, we obtain
a contradiction in both
cases, wich concludes
the proof.
\end{proof}

Let us define the
following syzygy:
$$\sigma := (t, 3, 1) + (1, 2, 1) - (1, 1, 1).$$

Then, the syzygy module
can be generated using $\sigma.$
\begin{lem}
	The union $U$ of $\{\sigma\}$ and of the principal syzygies is a GB
	of the syzygies.
\end{lem}
\begin{proof}
We show that any non-zero
syzygy is reducible by
an element of $U.$
It will be enough to 
conclude.
\begin{fact}
	If any non-zero
syzygy is reducible by
an element of $U$
then $U$ is a GB 
of the syzygy module.
\end{fact}
\begin{proof}
The order on $\MM \left( \mathscr{F}_3 \right)$
is isomorphic to $\NN.$
Let us then enumerate
increasingly the monomials
of $\MM \left( \mathscr{F}_3 \right)$:
$\alpha_1 < \alpha_2< \dots.$
For any element $a=\sum_i c_i \alpha_i$ of 
$\mathscr{F}_3,$
we associate the integer
$int(a)$:
$\sum_{i \textrm{ s.t. } c_i \neq 0} 2^i.$
Then any reduction
corresponds to a decrease
of $int.$
Then, eventually,
the $\LM$ of $a$
will be reduced,
and this is enough to conclure. 
\end{proof}

	Let $s$ be a non-zero syzygy.

	As the $g_i$'s are of the form $\alpha_i - \beta_i$, 
	we can define
	$$m := \min_{(l, \varepsilon_i, r) \mbox{ monomial of } s} l \beta_i r$$

	Since $\overline{s} = 0$, there is at least
	two monomials $s$, $(l_1, i_1, r_1)$ and $(l_2, i_2, r_2)$ of $s$, such that $l_1 \beta_{i_1} r_1 = l_2 \beta_{i_2} r_2 = m$.
	Then, $l_1 = l_2$ and $r_1 = r_2$ is not possible
	as it would mean that 
	$\beta_{i_1} = \beta_{i_2}$ and hence $i_1 = i_2$.

	If $\deg{l_1} \geq \deg{l_2 \beta_{i_2}}$ or $\deg{l_2} \geq \deg{l_1 \beta_{i_1}}$, then by
	symmetry, we can assume that $m$ can be written:
	$$m_1 \beta_{i_1} m_2 \beta_{i_2} m_3.$$
	Consequently:
	\begin{align*}
		(l_1, i_1, r_1) &= (m_1, i_1, m_2 \beta_{i_2} m_3), \\
		(l_2, i_2, r_2) &= (m_1 \beta_{i_1} m_2, i_2, m_3).
	\end{align*}
	Therefore, $s$ is reducible by the principal
	syzygy:
	$$m_1 (g_{i_1} (m_2, i_2, 1) - (1, i_1, m_2) g_{i_2}) m_3.$$

	Otherwise, two
	occurences of $\beta_{i_1}$ and $\beta_{i_2}$ in $m$
	have a non-empty intersection, which is not
	a complete identity.

	It is only possible if $\{i_1, i_2\} = \{1, 3\}$.
	By symmetry, let us
	assume that $i_1 = 1$ and $i_2 = 3$. In that cas,
	 $m$ can be written:
	$$m_1 ta m_2.$$
	Therefore:
	\begin{align*}
		(l_1, i_1, r_1) &= (m_1, \varepsilon_1, m_2), \\
		(l_2, i_2, r_2) &= (m_1 t, \varepsilon_3, m_2).
	\end{align*}
	Hence, $s$ est reducible by the syzygy:
	$$m_1 \sigma m_2.$$

	Consequently, any non-zero syzygy is divisible
	by an element of $U,$
	and this concludes the proof.
\end{proof}
